Formalizing Stack Safety as a Security Property

### 2022-10-10 submission5/noexceptions.tex
Responded to notes

We state our properties in terms of a generic machine subject to a few constraints,
but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
instruction, only {\tt jal} instructions which may be used for that purpose or potentially
other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
We therefore define our properties under the assumption that we can distinguish the
steps that represent these abstract operations from those that do not -- an assumption
that is valid in our testing environment (Section [TBD]) and that is already necessary
to correctly apply tag-based enforcement in practice (cite something about this.)

The set of special operations that are recognized in this way are termed
{\it security operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional information about its
security principals and which registers and memory they require to be secure, which evolves
dynamically through the execution of security operations.
Specifications may take the form of
predicates on future states: statements of the form,
``When control returns to me...'', or of relations on traces of future execution
(hyper-properties).

This approach is in contrast to approaches that give a safe-by-construction semantics
for various security operations and then prove secure compilation to the underlying
enforcement mechanism, in that our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to distinguish an
enforcement scheme that implements stack safety from one that does not.

@@

We must trust that our method of distinguishing security operations is reasonable; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If it recognizes the operations based on the dynamic behavior of
the system, similarly, those definitions must be realistic.

Otherwise, we do not assume that the code is reasonable or corresponds to any particular
calling convention or source construct. In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
and are drawn from the set (\(\WORDS\)) (by default, the set of 64-bit integers).
Registers are ranged over in general by \(\reg \in \REGS\), other than the
program counter \(\PCname\) and stack pointer \(\SP\).
They are divided into sets of caller-saved and callee-saved registers;
Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
appear in our examples and testing, along with the register names that will appear in
example code, and their default security class (explained below).

@@

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to values.
{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\).
The machine step function
\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by a list of security operations (order matters), defined below,
and an event.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by remaining in the
same state. Since failstops are not distinguished from normal execution,
our properties will suffer the limitations typical of
{\it termination insensitive} systems \cite{}.

@@

Security operations consist of calls (including tailcalls), returns,
allocations and deallocations. Calls and tailcalls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, the current
activation, and some number of pending ones. Each activation will have a {\it view}
of the system that maps each element to one of a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

These indicate, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
\(\unsealed\) for caller-saved, and \(\public\) otherwise.

A pending activation additional carries the addresses at which the program counter
and stack pointer should point in the event of a return.
We represent a {\it call stack} as a pair of the current activation's view
and a list of pending ones.
The initial call stack is \(\context_0 = (V_0, [ ~ ])\).
\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]

Every security operation manipulates the call stack via a function
\(Op : \Psi \rightarrow \MACHS \rightarrow \CONTEXTS \rightarrow \CONTEXTS\). We will introduce the
curried application of \(Op\) to each operation as we introduce them in detail.

From a machine state and a context stack, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with the same operations. Note that \(Op\) takes as its state argument
\(\mach\), the state before the step.

\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
            {\(\mathit{foldl} ~ (Op ~ \psi ~ \mach) ~ \context = \context'\)}
            {\((\mach,\context) \stepstounder{\psi :: \overline{\psi}} (\mach', \context'', \obs)\)}

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret in its arguments and allocates
an array of potentially sensitive data. It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data.

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    {\tt
      void main(int argc, int* argv) \{

@@

      ~ ~ int sensitive[4] = \{0,0,0,0\};

@@

      ~ ~ if (sensitive[0] == 42) \{

@@

      \}}
  \end{subfigure}
  \begin{subfigure}[b]{.45\columnwidth}
    \vspace{2\abovedisplayskip}

    {\tt
      0: addi sp,sp,-24

      4: sw a1,24(sp)

      8: sd zero,16(sp)

      12: sd zero,8(sp)

      16: jal 84,ra

      20: lw a4,4(sp)

      24: li a5,42

      28: bne a4,a5,L1

      32: lw a0,16(sp)

      36: jal 164,ra

      40: j L2:

      L1:

      44: jal 164,ra

      L2:

      48: addi sp,sp,24
    }
  \end{subfigure}
  \begin{subfigure}[b]{.5\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}. It might take an offset from the stack
pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, it might
instead return that value so that {\tt main} makes the call to {\tt publish}.

Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive[0]}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
And last, but certainly not least, the attacker might attempt to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.)

\begin{figure}
  \begin{subfigure}[b]{.5\columnwidth}
    {\tt
      100: addi sp,sp,-4

      104: sw ra,4(sp)

      108: lw a0,20(sp)

      112: jal 88,ra

      116: lw ra,4(sp)

      120: jalr ra
    }
    \caption{Leaking {\tt secret} directly}
  \end{subfigure}  
  \begin{subfigure}[b]{.45\columnwidth}
    {\tt
      100: lw a4,16(sp)

      104: sw a4,12(sp)

      108: jalr ra
    }
    \caption{Leaking {\tt secret} indirectly}
  \end{subfigure}

  \vspace{\belowdisplayskip}

  \begin{subfigure}[b]{.5\columnwidth}
    {\tt
      100: li a5,42

      104: sw a5,12(sp)

      108: jalr ra
    }
    \subcaption{Attacking {\tt sensitive}}
  \end{subfigure}
  \begin{subfigure}[b]{.45\columnwidth}
    {\tt
      100: addi ra,ra,20

      104: jalr ra
    }
    \subcaption{Attacking control flow}
  \end{subfigure}

@@

At the start of execution the program counter is 0 and the stack pointer -4, with code
proceeding upward and the stack growing down. The functions {\tt f} and {\tt publish}
are at addresses 100 and 200, respectively.

Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack
develop over several steps, with labels on the steps that allocate space and then make
a call. \(\downarrow \overline{\psi}\) in this diagram is a sideways
\(\xrightarrow{\overline{\psi}}\). Events are omitted.

The first step allocates a word for {\tt secret} and four for {\tt sensitive}. To
maintain a two-word alignment it leaves a gap. This will have the
effect of marking those bytes \(\object\), assuming they were previously
\(\unsealed\).

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l | l | l |}
      \hline
      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-4,4);\mathbf{alloc} ~ (-24,16)]\)} \\
      \hline
      1 & 4 & -24 & \(V_1 = V_0 \llbracket -24 \ldots -9, \mapsto \object \rrbracket\) \\
          & & & \(\llbracket -4 \ldots -1, \mapsto \object \rrbracket, \emplist\)\\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      4 & 16 & -24 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & 100 & -24 & \(V_2 = V_1 \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket,\) \\
      & & & \([(V_1, 20, -24)]\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      \(n\) & ? & ? & \(V_2,[(V_1, 20, -24)]\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      \(n+1\) & 20 & -24 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

@@

On a call, the formerly active principal's record is pushed onto the inactive list.
Its return target is the instruction immediately following that of the program
counter before the call, and the stack pointer target is the current stack pointer.
The callee's view is updated from the caller's such that the argument
registers are \(\public\) and any non-argument, caller-saved registers
are \(\unsealed\). All callee-save registers are always \(\sealed\) and
the program counter \(\public\), as they were initialized.

We suppose that, on step \(n+1\), {\tt f} returns; if it does not, then
{\tt main}'s secrets are safe as long as {\tt f} does not leak them directly.
On that return, we restore the topmost inactive principal, {\tt main}'s principal.

@@

But what if {\tt f}, does not return with \(\PCname = 20\) and \(\SP = -24\),
as would be the case in the fourth example attack? Then it has violated WBCF.
This property can be phrased in terms of a single step: for every transition
labeled as a return, the program counter and stack pointer match those recorded
in the top pending activation.

@@

In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses -24 through -9 and -4 through -1 (allocated as private data) and the callee-saved registers.
Note that the callee-saved registers may well have changed during the call -- but if
so, the callee is obligated to restore them to their prior value.

One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
in a way that is harmless -- or that is rendered harmless by the enforcement mechanism!
This is the case in a lazy micro-policy: the monitor permits the caller to write into the
callee's frame, but taints that write so that the callee cannot read it later without
triggering a failstop. In such a scenario we will consider a value {\it harmless}
if it could be replaced with any other value without changing the observable behavior
of the machine. That is, for a given state \(\mach\) and set of elements \(\components\),
we define a {\it variant} state that agrees with \(\mach\) on every element except those
in \(\components\). Those elements are said to be harmless in \(\mach\) if the variant
state produces an equivalent trace (see Section \ref{sec:events}.)

So, to define integrity, we need to know when a caller has been returned to;
this is the first state after the call in which the depth of the call stack
matches the state before the call. Integrity means that, at that return state,
any element that is \(\sealed\) under the callee's view is either restored
to its original value or is harmless.

@@

We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
means that its callee's behavior is dependent only on publicly visible data,
not the caller's state. As we saw in our examples, the callee's behavior is defined
both in terms of events that the callee produces during the call, but also in terms of
the changes it makes to the state.

@@

But, it is also possible that the callee could extract data from the caller that it does
not use to produce a visible event, but instead stashes for later or uses to influence
its return value -- which in turn can lead to future changes in behavior that do not
even appear to be the fault of the callee!
{\it Return-time confidentiality} means that any state element that changes in one variant
must either change in both, or be harmless.

@@

We model this as nearly dual to caller integrity: the callee's secrets are those
values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
at the state following the callee's return, any element that was \(\unsealed\) under the
callee's view either retains its original value or is harmless.

@@

At the most basic, if we protect nothing else, it is vital to protect the control flow of the
system by enforcing WBCF. None of the data-protection properties are meaningful if a
callee can simply ``return'' to an arbitrary instruction and proceed with the privileges
of its caller! And from a practical standpoint, control-flow attacks such as return-oriented
programming are among the most severe attacks involving the stack and the core motivation
for much existing work.

@@

\paragraph*{Variants and Harmless Values}

Two states are variants with respect to a set of elements, \(\components\),
if they agree on the value of every element not in \(\components\). Our
notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of vestigial elements.

@@

\begin{minipage}[t]{.6\columnwidth}
{\tt
  void main() \{

>>>>

We state our properties in terms of a generic machine subject to a few constraints,
but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
instruction, only {\tt jal} instructions which may be used for that purpose or potentially
other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
We therefore define our properties under the assumption that we can distinguish the
steps that represent these abstract operations from those that do not -- for instance,
by labeling instructions as calls, returns, etc. Such a labeling scheme appears in
our testing environment (Section [TBD]) and is already necessary
to correctly apply tag-based enforcement in practice (cite something about this.)

The set of special operations that are recognized in this way are termed
{\it security operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional context about its
security principals and which registers and memory they require to be secure.
This context evolves dynamically through the execution of security operations.
This notion of security may be phrased as
predicates on future states, e.g. assertions of the form
``When control returns to me...'', or relations on traces of future execution
(hyper-properties).

Unlike approaches that give a safe-by-construction semantics
for various security operations and then prove secure compilation to the underlying
enforcement mechanism, our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to judge whether
an enforcement scheme implements stack safety or not.

@@

We must trust that our method of distinguishing security operations is accurate; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If operations occur that are not recognized, those operations
might not be guaranteed to protect their principals -- for instance, an unlabeled call
might not protect the caller's data from the callee. On the flip side, applying an incorrect
label most likely means that the property becomes too strong to be enforced.

Otherwise, we do not assume that the code is reasonable or derived from any particular
calling convention or source construct. In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
and are drawn from the set \(\WORDS\).
Registers in the set \(\REGS\) are ranged over by \(\reg\), with special names given to the
program counter \(\PCname\) and stack pointer \(\SP\).
They are divided into sets of caller-saved and callee-saved registers;
Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
appear in our examples and testing, along with the register names that will appear in
example code, and their default security class (explained below).

@@

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to words.
{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
special silent event \(\tau\).
The machine step function
\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by an ordered list of security operations, defined below,
and an event.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by remaining in the
same state. Since failstops are not distinguished from normal execution,
our properties will suffer the limitations typical of
{\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.

@@

Security operations consist of calls (including tailcalls), returns,
allocations and deallocations. Calls and tailcalls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a{\it security classe}:
\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

A view indicates, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
\(\unsealed\) for caller-saved, and \(\public\) otherwise.

A pending activation also carries the addresses to which the program counter
and stack pointer should point in the event of a return. We represent the call
stack with a list of such triples. A {\it context} is a pair of the current activation's view
and a list of pending ones.
The initial context is \(\context_0 = (V_0, [ ~ ])\).
\[(V, \addr_{ret}, \addr_{sp}) \in \mathit{PEND} ::= \mathit{VIEW \times \WORDS \times \WORDS}\]
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ PEND}\]

Every security operation manipulates the call  stack via a function
\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).

From a machine state and a context, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with the same operations. Note that \(Op\) takes as its state argument
\(\mach\), the state before the step.

\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
            {\(\mathit{foldl} ~ (Op ~ \mach) ~ \context ~ \overline{\psi} = \context'\)}
            {\((\mach,\context) \stepstounder{\overline{\psi}} (\mach', \context'' \obs)\)}

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret in its arguments and initializes a variable to contain
potentially sensitive data.
It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not -- instead it will publish the results of {\tt f}.

\begin{figure}
  \begin{subfigure}{\columnwidth}
    {\tt
      void main(int argc, int* argv) \{

@@

      ~ ~ int sensitive = 0;

@@

      ~ ~ if (sensitive == 42) \{

@@

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a1,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal 100,ra}{\(\mathbf{call} ~ 100 ~ \emplist ~ \emplist\)}
      \labeledrow{20:}{sw r0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
      \labeledrow{52:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0}\} ~ \emplist\)}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.
In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
instead return that value so that {\tt main} makes the call to {\tt publish},
as in Figure \ref{subfig:indirect}.

Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(Figure \ref{subfig:integrity})!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
And in Figure \ref{subfig:WBCF}, the attacker attempts to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.)

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi sp,sp,-8}{\(\mathbf{alloc} ~ (-8,8)\)}
      \labeledrow{104:}{sd ra,0(sp)}{}
      \labeledrow{108:}{lw a0,12(sp)}{}
      \labeledrow{112:}{jal 200,ra}{\(\mathbf{call} ~ 200 ~ \{\mathtt{a0} ~ \emplist\}\)}
      \labeledrow{116:}{lw ra,0(sp)}{}
      \labeledrow{120:}{addi sp,sp,8}{\(\mathbf{dealloc} ~ (0,8)\)}
      \labeledrow{124:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} directly}
    \label{subfig:direct}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a0,12(sp)}{}
      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} indirectly}
    \label{subfig:indirect}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{li a5,42}{}
      \labeledrow{104:}{sw a5,4(sp)}{}
      \labeledrow{108:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking {\tt sensitive}}
    \label{subfig:integrity}
  \end{subfigure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi ra,ra,20}{}
      \labeledrow{104:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking control flow}
    \label{subfig:WBCF}
  \end{subfigure}

@@

At the start of execution the program counter is 0 and the stack pointer -4, with the
stack growing down. The functions {\tt f} and {\tt publish}
are at addresses 100 and 200, respectively.

Figure \ref{fig:exec1} shows how the program counter, stack pointer, and context
develop over several steps, with labels on the steps that allocate space and then make
a call. \(\downarrow \overline{\psi}\) in this diagram is a sideways
\(\xrightarrow{\overline{\psi}}\). Events are omitted.

The first step  allocates words for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
as two words for the return address. This will have the
effect of marking those bytes \(\object\), assuming they were previously
\(\unsealed\).

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l | l || l |}
      \hline
      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
      \hline
      1 & 4 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      4 & 16 & -20 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & 100 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[(V_1, 20, -20)]\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      \(n\) & ? & ? & \(V_2,[(V_1, 20, -20)]\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      \(n+1\) & 20 & -20 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

@@

On a call, the formerly active principal's record is pushed onto the inactive list.
Its return target is the return address of the call, 
and the stack pointer target is the stack pointer at the moment of call.
The callee's view is updated from the caller's such that all \(\object\) locations
become \(\sealed\). (In more complex settings, objects that are intentionally passed
to the callee will not get sealed, but for now we assume no sharing of memory.)
For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
in this case {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
registers are mapped to \(\unsealed\). All callee-save registers are always \(\sealed\) and
the program counter \(\public\), as they were initialized.

On step \(n\), {\tt f} returns, and we restore the topmost inactive principal,
{\tt main}'s principal.

@@

But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or \(\SP \neq -24\)),
as would be the case in attack (d)? Then it has violated the WBCF property.
This property can be phrased in terms of a single step: for every transition
labeled as a return, the program counter and stack pointer at the next state
match those recorded in the top pending activation. The table in Figure \ref{fig:exec1}
assumes that WBCF is followed.

@@

In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses -20 through -1 (allocated as private data) and callee-saved registers such as
the stack pointer and return address.
Note that the callee-saved registers may well have changed during the call -- but if
so, the callee is obligated to restore them to their prior value.

One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
in a way that is harmless -- or that is rendered harmless by the enforcement mechanism!
This is the case in a lazy micro-policy: the monitor permits the caller to write into the
callee's frame, but taints that write so that the callee cannot read it later without
triggering a failstop. In such a scenario we will consider a value {\it irrelevant}
if it could be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a given state \(\mach\) and set of elements \(\components\),
we define a {\em \(\components\)-variant} state to be one agrees with \(\mach\) on every element except those
in \(\components\). The elements of \(\components\) are said to be irrelevant in \(\mach\) if every
variant state produces an equivalent trace (see Section \ref{sec:events}.)

So, to define integrity, we need to know when a caller has been returned to;
this is the first state after the call in which the depth of the call stack
matches the state before the call. Integrity holds if, at that return state,
any element that is \(\sealed\) under the callee's view is either restored
to its original value or is irrelevant.

Returning to our example: at step 4, when {\tt main} makes its call to {\tt f},
the call stack in the context is empty. It makes its call, and at step 5, {\tt main}'s
frame, including {\tt secret} and {\tt sensitive}, is mapped to \(\sealed\) from the
perspective of {\tt f}. So, the next time the call stack is empty (implicitly, when
we have returned to {\tt main}), elements of that frame must either match their
values at step 4, or be irrelevant to future computation.

@@

\apt{Remainder of this section is on the right track, but needs some expansion,
  and to be tied back more clearly and consistently to the examples. I started
  word-smithing this first para...}
We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
means that its callee's behavior is dependent only on publicly visible data,
not the caller's private state. As we saw in the examples, we must consider both the observable events
that the callee produces during the call and the changes that the callee makes to the state that might
affect the caller after the callee returns.

@@

But, in Figure \ref{subfig:indirect}, we also saw an attacker that exfiltrated the secret
by reading it and then returning it, in a context where the caller would output the returned
value. In situations like this, we care about {\it return-time confidentiality}.
In that case, we would identify that in our original execution, {\tt a0} changed between
the call and return, being set to the value of {\tt secret}. But in a variant state,
{\tt secret} might have a different value, and so will {\tt a0} at the return.
This difference, found by comparing the initial and final states of both executions,
indicates that {\tt a0} is corrupted. Return-time confidentiality means that
an element that is corrupted in this way must be irrelevant.

@@

We model this as nearly dual to caller integrity: the callee's secrets are those
values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
at the state following the callee's return, any element that was \(\unsealed\) under the
callee's view either retains its original value or is irrelevant.

@@

At the most basic, if we protect nothing else, it is vital to protect the control flow of the
system by enforcing WBCF or a similar property. None of the data-protection properties are
meaningful if a callee can simply ``return'' to an arbitrary instruction and proceed with
the privileges of its caller! And from a practical standpoint, control-flow attacks such
as return-oriented programming are among the most severe attacks involving the stack and
the core motivation for much existing work.

That said, many systems will not enforce WBCF, as it is insufficient to capture the
control flow of features such as exceptions. So, the data-protection properties are
technically orthogonal to WBCF. It remains the case that an appropriate control-flow
property (extended with relevant control-flow features) is a vital base to any stack safety
enforcement.

Once control-flow is protected, we can think about integrity and confidentiality.
Integrity is perhaps a more obviously useful property: major proposed formalizations
of stack safety have omitted confidentiality in favor of combining WBCF and integrity
\cite{SkorstengaardSTKJFP}. This is because integrity is directly tied to compiler correctness
in a way that confidentiality is not: compilers may optimize based on assumptions
about the invariance of a caller's frame over a call, and without integrity such optimizations
are in fact unsound.

Finally, at the strongest end of the hierarchy, we might enforce safe control flow,
integrity, and confidentiality of both kinds.

@@

\paragraph*{Variants and Irrelevant Values}

Two states are variants with respect to a set of elements, \(\components\),
if they agree on the value of every element not in \(\components\). Our
notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements.

@@

{\tt
  void main() \{
